Scalar Items in Embedded Position: {A}n Experimental Revisit

### 2013-01-21 text/paper/paper3.tex
paper3 - up to general discussion

\paragraph{General Idea.} In order to test the availability and
preferences of different readings of \as- and \es-sentences in German, 
we were looking for a way to unambiguously map responses from a picture
verification task to specific readings. We were particularly
interested in the question whether local readings would be observed at
all, and whether their presence or absence would be modulated by
a contrastive accent on the scalar item. Moreover, we were  
interested in obtaining information about the relative preferences 
of the different readings. In addition, we wanted
to minimize effects of graphical differences between the picture
materials. To achieve this, we designed an experiment employing a
modified version of picture verification, namely the
\mymark{incremental verification task} (\acro{ivt}, SEE CONROY
????). The general idea is that subjects are requested to judge
sentence material based on pictures that do not necessarily contain
all the information necessary to judge a certain reading true or
false. In that case, participants can demand that more information is
revealed. Participants are instructed to make a truth-value judgment
as soon as they are able to.\dn{need to mention somewhere that trial
  ends with each TV judgment}

\paragraph{Target Conditions.} In the present study, participants were 
presented pictures
%, like those in
%Figures~\ref{fig:InitialPicture}, \ref{fig:exseqAS} or
%\ref{fig:exseqES}, 
depicting a set of four identical central elements
(e.g., letters), which could be connected to surrounding elements
(e.g., triangles). Initially, any potential connections between
central and surrounding elements were covered by dark gray color (see
Figure~\ref{fig:InitialPicture}).
%
\begin{figure}
  \centering
      \fbox{\includegraphics[width=5cm]{../pictures/paper/ae_01_1.pdf}}
      \caption{Initial picture presented to participants where all
        possible connections from each letter to its surrounding
        triangles are covered.}
  \label{fig:InitialPicture}
\end{figure}
%
Sentences to be judged 
%were in German (see (\ref{ex:as}) and (\ref{ex:es}) below) and 
were presented auditorily, and participants were
asked to listen to a given sentence and then uncover the potential
links step by step until they felt able to give a truth value
judgment. Three options were available for participants at each step:
(i), demand more information, (ii), judge the sentence as true, and
(iii), judge the sentence as false. Importantly, each of the three
potentially available readings corresponded to a specific step in the
uncovering process where the truth value of that reading (and only of
that reading) could be assessed for the first time, which we will
refer to as {\it critical position}. The location in the sequence and 
the corresponding truth-value judgment differed between \as- and 
\es-sentences, as described presently. (This is, again, because of the 
logical dependencies between readings.)

Consider the \as-sentence in (\ref{ex:as}).
\begin{exe}
\ex \gll \mymark{Alle} diese Briefe sind mit \mymark{einigen} ihrer Dreiecke
  verbunden.\label{ex:as}\\
All these letters are with some their triangles connected.\\
\trans All of these letters are connected to some of their triangles.
\end{exe}
Figure~\ref{fig:exseqAS} shows the three corresponding critical
positions for this sentence in the order of the uncovering
process. First, the situation in Figure~\ref{fig:exseqAS1} becomes
available, which is true under a literal reading, while the local and
global readings cannot be evaluated yet. Second, in
Figure~\ref{fig:exseqAS2}, the literal reading is still true, but now
the global reading can be additionally confirmed by a \emph{true}
response, as all the connections of one of the letters are now
uncovered, and it is now visible that one of the letters is not
connected to all of its triangles. Finally, decisions concerning the
local reading are possible as soon as all connections have been
uncovered. In this case, the critical position shown in
Figure~\ref{fig:exseqAS3} is incompatible with a local reading: One of
the central elements is linked to all of its surrounding elements,
thus yielding to \emph{false} answers if participants indeed adopted
the local reading. Note that associating the final position with
\emph{false} answers allows us to exclude both literal and global
readings, which correspond to a \emph{true} answer at this
position. In sum, \emph{true} or \emph{false} answers on particular
positions in the incrementally revealed picture sequence can be mapped
uniquely to candidate reading. All other \emph{true} or \emph{false}
answer were counted as errors. The mapping between responses and
readings is summarized in Table~\ref{tab:mappingAS}.

@@

\es-sentences, due to the non-linear relationship of candidate
readings described in Section~\ref{sec:get-know-your}, require a
slightly different sequence of unfolding which yields a different
order in which truth-value judgments can be made.
%
\begin{figure}[ht]
	\centering
	\subfloat[][Step 1]{ 
		\fbox{\includegraphics[width=3cm]{../pictures/paper/ge_3_v_l.pdf}}
	    \label{fig:subfig1}
	}
	\subfloat[][Step 2]{
		\fbox{\includegraphics[width=3cm]{../pictures/paper/ge_4_v_l.pdf}}
	    \label{fig:subfig2}
	}
	\subfloat[][Step 3]{ 
		\fbox{\includegraphics[width=3cm]{../pictures/paper/ge_6_v_l.pdf}}
	    \label{fig:subfig1}
	}
	\caption[]{Example sequence for \es-sentences}
	\label{fig:exseqES}
\end{figure}
%
For a sentence like (\ref{ex:es}) we obtain an unambiguous
mapping between truth values and readings with respect to the sequence
in \ref{fig:exseqES}.
\begin{exe}
\ex \label{ex:es} \gll \mymark{Genau} eine der Glocken ist mit
  \mymark{einigen} ihrer Halbkreise verbunden.\\ 
  Exactly one of-the bells is with some its semicircles connected.\\
  \trans Exactly one bell is connected to some of its semicircless.
\end{exe}
At the first critical position shown in Figure~\ref{fig:subfig1}, the
global reading corresponds to a \emph{false} response, as one of the
bells is already linked to all of its semicircles. The literal reading
can be evaluated in Figure~\ref{fig:subfig2}, where it would trigger a
\emph{false} response, as there are two bells that are connected to at
least some of their surrounding elements. Finally, confirming
\ref{fig:subfig1} by a \emph{true} response corresponds to the local
reading, as one of the bells is linked to some but not all of its
semicircles, while all of the other bells are linked to all of
them. Note that again, global and literal readings differ from local
readings with respect to their truth values. Mappings between
responses and readings are summarized in Table \ref{tab:mappingES}.

\paragraph{Positional controls.} In order to make sure that subjects
understood the task, i.e., gave truth-value judgements neither sooner
not later than at the first possible position in a sequence, we
included a set of control conditions. These also controlled for
response biases that are independent of the sentence meaning, but only
depend on the experimental procedure or on the picture materials. For
each of the three readings in the \as- and \es-conditions, we
constructed an unambiguous control sentence as in
(\ref{bsp:controls-as}) and (\ref{bsp:controls-es}), requiring the
same judgement at the identical position in the same sequence used for
the respective targets. For instance, example
(\ref{bsp:controls-as-1}) requires a \emph{true}-response analogous to
the \as-sentence in (\ref{ex:as}) under its literal reading,
(\ref{bsp:controls-as-2}) corresponds to the \as-sentence under its
global reading and (\ref{bsp:controls-as-3}) corresponds to its local
reading. With regard to the \es-sentences, controls like
(\ref{bsp:controls-es-1}) correspond to the global reading,
(\ref{bsp:controls-es-2}) to the literal and (\ref{bsp:controls-es-3})
to the local reading. If, independently of sentence meaning, there was
any bias to respond in a certain way at any point in the sequences,
such as to preferably unravel the whole picture, this should affect
control sentences to the same degree as it affects target conditions.

@@

\paragraph{Preference-related controls.} In order to control for
artificial effects of presentation order on the availability of
readings, we included a second type of control conditions. In
particular, we tested (i), whether the order of given readings affects
preferences for ambiguous sentences and (ii), whether prosodic
information can, in principle, shift reading preferences in the
present task. Note that for our target sentences, the logical
entailment relations between readings always require local readings to
be evaluated at the end of a sequence. It could thus in principle be
possible that no subject ever reaches this point, because they gave
truth-value judgements earlier, thereby ending the trial. In that case
it would be unclear if these decisions have been affected by a general
inavailability of local readings or by the fact that one of the
earlier presented readings is the preferred one. By including globally
ambiguous structures with a known preference over available readings,
we thus intended to test whether participants in our task occasionally
choose dispreferred readings, even if they were available only at a critical
position following the preferred readings.

@@

\end{exe}
In late-closure sentences, a modifier such as a relative clause or a 
prepositional phrase can be attached to one of two
preceding hosts. For instance, in (\ref{bsp:target-related}), the
prepositional phrase \emph{with suns} can be attached to the preceding
noun \emph{squares}, resulting in the so-called \emph{late-closure} or
\emph{low-attachment} reading ((\ref{bsp:target-related-LC}). Alternatively, 
the \acro{pp} can be attached to the whole conjoined \acro{np}
\emph{circles and squares}, corresponding to the \emph{early-closure}
or \emph{high-attachment} reading (\ref{bsp:target-related-EC}).
Late-closure readings have been observed to be preferred over the early 
closure reading for sentences comparable to those in (\ref{bsp:target-related})
\citep[e.g.][]{Frazier87,Frazier82}.

For each sentence, two sequences were designed. In the first one, the
critical position associated with the dispreferred early-closure
reading preceded the critical position for the preferred late-closure
reading. In this case, a partly covered picture as in
Figure~\ref{fig:exec1} preceded a picture as in \ref{fig:exec2}.
%
\begin{figure}[ht]
	\centering
	\subfloat[][Step 1]{ 
		\fbox{\includegraphics[width=5.5cm]{../pictures/paper/ec_01_3.pdf}}
	    \label{fig:exec1}
	}
	\subfloat[][Step 2]{
		\fbox{\includegraphics[width=5.5cm]{../pictures/paper/ec_01_5.pdf}}
	    \label{fig:exec2}
	}
	\caption[]{Critical steps of a sequence for sentence
          (\ref{bsp:target-related}) where the early-closure reading
          (\ref{bsp:target-related-EC}) can be judged first.}
	\label{fig:exec}
\end{figure}
%
If participants adopted an early-closure reading as in
(\ref{bsp:target-related-EC}), a \emph{false}-judgment would be
expected at the position illustrated in \ref{fig:exec1}, as none of
the circles connected to the letter contain any suns. Under the
late-closure reading in (\ref{bsp:target-related-EC}), in contrast, a
\emph{true}-response would be expected no sooner than in the step
shown in Figure~\ref{fig:exec2}. The order of possible judgements is
reversed in the second kind of sequence including pictures like
\ref{fig:exlc1} followed by pictures like \ref{fig:exlc2}. Here, the
late-closure reading in (\ref{bsp:target-related-LC}) can be judged
first requiring a \emph{true}-response on \ref{fig:exlc1}. The
early-closure reading in (\ref{bsp:target-related-EC}) can only be
judged later in the sequence. When we reach \ref{fig:exlc2} a
\emph{false}-response is expected under an
early-closure/high-attachment reading.

@@

\begin{enumerate}[1:]
\item a higher higher f0 maximum or a higher pitch range (difference 
between minimal and maximal f0 values), and
\item a longer duration of the accented element as opposed to its 
non-accented counterpart
\end{enumerate}

We therefore recorded sentences by a trained speaker familiar with the 
concept of contrastive focus and instructed her to produce contrast as 
described in 1 and 2. Both \as-sentences and \es-sentences
were recorded in two versions each. In the accented version, a pitch 
accent was placed on {\it einige}. The second version had neutral prosody. 
If accentuation is the driving force for local readings, we would expect 
a higher proportion of local readings in the former than in the latter 
version of the sentences.

Preference-related controls also differed prosodically. Modifier attachment 
ambiguities comparable to the examples in (\ref{bsp:target-related}) and 
(\ref{bsp:target-related2}) have been shown to be sensitive to differences 
in prosodic phrasing. Though there has been discussion as to whether 
speakers reliably \textit{produce} disambiguating cues in such constructions 
\citep[e.g.][]{Allbritton96,Kraljik05, Snedeker03, Schafer00}, overt prosodic 
boundaries are generally acknowledged to guide interpretive processes of Late 
Closure sentences in \textit{comprehension} \citep[e.g.][]{Steinhauer99, Augurzky06}. 
Specifically, it has been shown that prosodically separating a modifier from 
the directly preceding material supports an early closure reading, whereas 
separating the two potential attachment sites is supportive of a late closure 
reading. Thus, a phrase boundary between {\it squares} and {\it with} in
(\ref{bsp:target-related}) corresponds to a high-attachment reading, whereas
a prosodic phrase boundary following {\it circles} corresponds to a low-attachment
reading.

To test whether prosodic information is taken into account in the \acro{itv} task, we
presented sentences in both of these prosodic variants. In addition, in order to test 
prosody-independent preferences, we also included a neutral version without any 
pronounced boundary. Further, we also presented 
unambiguous sentences corresponding to the early-closure reading. As in the case 
of the \as- and \es-sentences, the unambiguous counterparts served as a baseline to
control for response biases that are independent of sentence meaning.

@@

We constructed a set of 15 items for \as- and \es-sentences,
respectively, analogous to the examples in \ref{ex:as} and
\ref{ex:es}. Sentences were introduced by a subject {\small
  \acro{DP}}, including the quantifiers {\it alle diese} ("all of
these") or {\it genau eine(r) der} ("exactly one of these"), as well
as the head noun which denoted different icons, such as letters or
bells. Subjects were followed by an auxiliary, and a {\small
  \acro{PP}} containing {\it einige}, a possessive pronoun, and a noun
denoting a geometrical object, e.g. {\it einigen seiner
  Dreiecke/Quadrate} ("some of its triangles/squares"). The possessive
pronoun was used in order to fix relative quantifier scope to surface
scope, thus ensuring that {\it einige} has embedded status. The last
word was the main verb {\it verbunden} ("connected").  Each sentence
was recorded in a stressed and an unstressed version of scalar {\it
  einige}. As described above, three control
sentences were created for each experimental item, corresponding to
the critical positions in the course of uncovering the accompanying
sequences.  The positional controls were recorded with
neutral prosody.  Items were evenly distributed across 5 lists using a
Latin square design.

@@

For all target sentences (\as and \es), the determiner \emph{einige} was
produced with a contrastive pitch accent as well as with a neutral
accent. Contrastive accents in German are realized by an increase in f0 (range)
as well as by a longer duration (see {sec:design} above). A set of 15 
experimental items was recorded for each condition (\as vs.~\es, \emph{accented} 
vs.~\emph{unaccented}), resulting in a total number of 60 target sentences.

@@

Note that differences between Regions 1, 2, 4 and 7 can be expected
due to lexical differences between the \as and \es-conditions. As
preference-related fillers did not differ with respect to their lexical
properties, we carried out word-by-word analyses for these
conditions. Durational values included the respective word plus any
following silent interval. Note that we did not include the
disambiguated fillers in these analyses as they involved very
different sentence types (i.e., constructions involving prepositional
phrases vs.~relative clauses).

@@

For the durational analyses, constituents were automatically labeled
by the \emph{Aligner} tool \citet{Rapp98},and the
obtained values were manually corrected afterwards. For the targets,
two-factorial \acros{anova} with the factors \acro{Quantifier} (all
vs. exactly one) and \acro{Prosody} (accented vs.~unaccented) were
carried out. For the preference-related fillers, we carried out
one-factorial \acros{anova} with the factor \acro{Prosody} (early
boundary, late boundary, neutral prosody). F0 values were extracted by
means of special Praat scripts
(\url{http://www.fon.hum.uva.nl/praat/}). For the present analyses,
differences between minimal and maximal F0 values for each region or
word were calculated. Again, two-factorial \acros{anova} were carried
out for statistical comparison of the target sentences, and one-factorial
\acros{anova} were carried out for preference-related controls.

The details of these analyses are found in
Appendix~\ref{sec:audit-sent-mater}. \dn{Petras Kommentar: Vielleicht sollten wir hier jeweils 1 Figure für akzentuiert vs. nicht und boundary vs. nicht reinbringen. Fabian meint zurecht, dass man nicht einfach die Ergebnisse nur in den Appendix packen kann. Es wären dann halt 4 Figures, da wir ja f0 und duration ansehen, und wir würden nur die experimentell interessanten veranschaulichen und die Statistik dann im Appendix lassen - was meint Ihr?}

In sum, our speaker reliably
produced (i) differences in accent realization for the target
sentences and (ii) the expected boundary realizations for the
preference-related controls. Whereas accented elements clearly
differed from their unaccented counterpart by showing an increase in
duration and F0 range, prosodic boundaries for our preference-related
fillers were realized by pre-final lengthening (i.e. an increase in F0
and duration at the position preceding the boundary).

@@

For the preference-related controls, we used two types of
sequences. First, a sequence in which the crirical position for the
(preferred) Late Closure reading was presented preceding the critical
position for the Early Closure reading ({\it LC sequence}), and
second, a sequence in which the readings were presented in reversed
order ({\it EC sequence}. In the LC sequences, seven pictures were
presented, with the LC reading given at the third picture, and the EC
reading at the sixth picture. EC sequences consisted of five pictures,
in which the critical positions were realized at the third picture (EC
reading) and at the fifth picture (LC reading).

@@

In the following, we will report the results for targets and positional
controls separately. After providing the descriptive results, we present
statistical analyses, which were carried out in the following manner.
We computed log-liner models (see Scheepers 2003) in two variants (by 
item and by participant). In the first, the factor {\it Item} was 
included, and in the second analysis, the factor {\it Participants} 
was included. Inclusion of these two factors allowed us to test whether 
the distribution of participants' choices was identical accros items and 
participants. We report log-likelihood ratio Chi-squares ($LRCS_1$ 
and $LRCS_2$), degrees of freedom ($df_1$, $df_2$) and significance 
levels ($p_1$ and $p_2$). The relevant experimental
factors are provided in the specific subsections below.

\paragraph{Target conditions and positional controls}
Performance on positional controls was high (92\% correct
responses on average), indicating that subjects understood the task,
giving truth-value judgements at the ``right'' moment in a sequence
without being influenced in general by inessential features of the
picture material or showing general response biases.

@@

      \begin{axis}[ybar, 
                  enlarge x limits=0.5,
                  enlarge y limits=0.25,
                  legend style={at={(0.5,-0.15)}, anchor=north, legend columns=-1},
        ylabel={\% of answers}, symbolic x
        coords={neutral,accented}, xtick=data, nodes
        near coords, nodes near coords align={vertical},x=65, bar
        width=3mm, ymin = 8, ymax = 73 ]

@@

\subsection{Discussion}
\label{sec:interpretation}

>>>>

\paragraph{General Idea.} In order to test the availability and
preferences of different readings of \as- and \es-sentences in German, 
we were looking for a way to unambiguously map responses from a picture
verification task to specific readings. We were particularly
interested in the question whether local readings would be observed at
all, and whether their presence or absence would be modulated by
a contrastive accent on the scalar item. Moreover, we were  
interested in obtaining information about the relative preferences 
of the different readings. In addition, we wanted
to minimize all effects of alleged typicality of pictures. To achieve this, we designed an experiment employing a
modified version of picture verification, namely the
\mymark{incremental verification task} (\acro{ivt}, SEE CONROY
????). The general idea is that subjects are requested to judge
sentence material based on pictures that do not necessarily contain
all the information necessary to judge a certain reading true or
false. In that case, participants can demand that more information be
revealed. Participants are instructed to make a truth-value judgment
as soon as they are able to. When they do, the trial ends.

\paragraph{Target Conditions.} In the present study, participants were
presented with pictures depicting a set of four identical central
elements (e.g., letters), which could be connected to surrounding
elements (e.g., triangles). Initially, any potential connections
between central and surrounding elements were covered by dark gray
color (see Figure~\ref{fig:InitialPicture}).
%
\begin{figure}
  \centering
      \fbox{\includegraphics[width=5cm]{../pictures/paper/ae_01_1.pdf}}
      \caption{Initial picture presented to participants where all
        possible connections from each letter to its surrounding
        triangles are covered.}
  \label{fig:InitialPicture}
\end{figure}
%
Sentences to be judged 
%were in German (see (\ref{ex:as}) and (\ref{ex:es}) below) and 
were presented auditorily, and participants were asked to listen to a
given sentence and then uncover the potential links step by step until
they felt able to give a truth value judgment. Three options were
available for participants at each step: (i) demand more information,
(ii) judge the sentence as true, and (iii) judge the sentence as
false. Trials ended with a truth-value judgement, and a judgement was
required to end the trial when all connections had been
revealed. Importantly, each of the three potentially available
readings corresponded to a specific step in the uncovering process
where the truth value of that reading (and only of that reading) could
be assessed for the first time. We refer to this step in the sequence
as the {\it critical position} of a given reading. The critical
position and the corresponding truth-value judgment differed between
\as- and \es-sentences, as described presently. (This is, partly,
because of the logical dependencies between readings, and, partly,
also to rule out positional biases.)

Consider the German \as-sentence in (\ref{ex:as}), included in our study.
\begin{exe}
\ex \gll \mymark{Alle} diese Briefe sind mit \mymark{einigen} ihrer Dreiecke
  verbunden.\label{ex:as}\\
All these letters are with some their triangles connected.\\
\trans All of these letters are connected to some of their triangles.
\end{exe}
Figure~\ref{fig:exseqAS} shows the three corresponding critical
positions for this sentence in the order of the uncovering
process. First, the situation in Figure~\ref{fig:exseqAS1} becomes
available, which is true under a literal reading, while the local and
global readings cannot be evaluated yet. Second, in
Figure~\ref{fig:exseqAS2}, the literal reading is still true, but now
the global reading can be additionally confirmed by a \emph{true}
response, as all the connections of one of the letters are now
uncovered, and it is now visible that one of the letters is not
connected to all of its triangles. Finally, decisions concerning the
local reading are possible as soon as all connections have been
uncovered. In this case, the critical position shown in
Figure~\ref{fig:exseqAS3} is incompatible with a local reading: One of
the central elements is linked to all of its surrounding elements,
thus yielding to \emph{false} answers if participants indeed adopted
the local reading. Note that associating the final position with
\emph{false} answers allows us to exclude both literal and global
readings, which correspond to a \emph{true} answer at this
position. In sum, \emph{true} or \emph{false} answers on particular
positions in the incrementally revealed picture sequence can be mapped
uniquely to candidate reading. All other \emph{true} or \emph{false}
answer were counted as errors. The mapping between responses and
readings is summarized in Table~\ref{tab:mappingAS}.

@@

\es-sentences, due to the non-linear relationship of candidate
readings described in Section~\ref{sec:get-know-your}, require a
slightly different sequence of unfolding which yields a different
order in which truth-value judgments can be made.
%
\begin{figure}[ht]
	\centering
	\subfloat[][Step 1]{ 
		\fbox{\includegraphics[width=3cm]{../pictures/paper/ge_3_v_l.pdf}}
	    \label{fig:subfig1}
	}
	\subfloat[][Step 2]{
		\fbox{\includegraphics[width=3cm]{../pictures/paper/ge_4_v_l.pdf}}
	    \label{fig:subfig2}
	}
	\subfloat[][Step 3]{ 
		\fbox{\includegraphics[width=3cm]{../pictures/paper/ge_6_v_l.pdf}}
	    \label{fig:subfig1}
	}
	\caption[]{Example sequence for \es-sentences}
	\label{fig:exseqES}
\end{figure}
%
For a sentence like (\ref{ex:es}) we obtain an unambiguous
mapping between truth values and readings with respect to the sequence
in \ref{fig:exseqES}.
\begin{exe}
\ex \label{ex:es} \gll \mymark{Genau} eine der Glocken ist mit
  \mymark{einigen} ihrer Halbkreise verbunden.\\ 
  Exactly one of-the bells is with some its semicircles connected.\\
  \trans Exactly one bell is connected to some of its semicircless.
\end{exe}
At the first critical position shown in Figure~\ref{fig:subfig1}, the
global reading corresponds to a \emph{false} response, as one of the
bells is already linked to all of its semicircles. The literal reading
can be evaluated at the critical position shown in
Figure~\ref{fig:subfig2}, where it would trigger a \emph{false}
response, as there are two bells that are connected to at least some
of their surrounding elements. Finally, confirming \ref{fig:subfig1}
by a \emph{true} response corresponds to the local reading, as one of
the bells is linked to some but not all of its semicircles, while all
of the other bells are linked to all of them. Note that again, global
and literal readings differ from local readings with respect to their
truth values. Mappings between responses and readings are summarized
in Table \ref{tab:mappingES}.

\paragraph{Positional controls.} In order to make sure that subjects
understood the task, i.e., gave truth-value judgements neither sooner
not later than at the first possible position in a sequence, we
included a set of control conditions. These also controlled for
response biases. For each of the three readings in the \as- and
\es-conditions, we constructed an unambiguous control sentence as in
(\ref{bsp:controls-as}) and (\ref{bsp:controls-es}), requiring the
same judgement at the identical position in the same sequence used for
the respective targets. For instance, example
(\ref{bsp:controls-as-1}) requires a \emph{true}-response analogous to
the \as-sentence in (\ref{ex:as}) under its literal reading,
(\ref{bsp:controls-as-2}) corresponds to the \as-sentence under its
global reading and (\ref{bsp:controls-as-3}) corresponds to its local
reading. With regard to the \es-sentences, controls like
(\ref{bsp:controls-es-1}) correspond to the global reading,
(\ref{bsp:controls-es-2}) to the literal and (\ref{bsp:controls-es-3})
to the local reading. If, independently of sentence meaning, there was
any bias to respond in a certain way at any point in the sequences,
such as to preferably unravel the whole picture, this should affect
control sentences to the same degree as it affects target conditions.

@@

\paragraph{Preference-related controls.} In order to test whether the
order of critical positions for different readings within a sequence
had an influence on responses, we included a second type of control
conditions. These conditions also controlled whether prosodic
information can, in principle, shift reading preferences in the
present task. Note that for our target sentences, the logical
entailment relations between readings always require local readings to
be evaluated at the end of a sequence. It could thus in principle be
possible that no subject ever reaches this point, because they gave
truth-value judgements earlier, thereby ending the trial. In that case
it would be unclear if these decisions have been affected by a general
unavailability of local readings or by the fact that one of the
earlier presented readings is the preferred one. By including globally
ambiguous structures with a known preference over available readings,
we thus intended to test whether participants in our task occasionally
choose dispreferred readings, even if they were available only at a
critical position following the preferred readings.

@@

\end{exe}
In late-closure sentences, a modifier such as a relative clause or a
prepositional phrase can be attached to one of two preceding
hosts. For instance, in (\ref{bsp:target-related}), the prepositional
phrase \emph{with suns} can be attached to the preceding noun
\emph{squares}, resulting in the so-called \emph{late-closure} or
\emph{low-attachment} reading
((\ref{bsp:target-related-LC}). Alternatively, the \acro{pp} can be
attached to the whole conjoined \acro{np} \emph{circles and squares},
corresponding to the \emph{early-closure} or \emph{high-attachment}
reading (\ref{bsp:target-related-EC}).  Late-closure readings have
been observed to be preferred over the early-closure reading for
sentences comparable to those in (\ref{bsp:target-related})
\citep[e.g.][]{Frazier87,Frazier82}.

For each sentence, two sequences were designed. In the first one, the
critical position associated with the dispreferred early-closure
reading preceded the critical position for the preferred late-closure
reading. In this case, a partly covered picture as in
Figure~\ref{fig:exec1} preceded a picture as in \ref{fig:exec2}.
%
\begin{figure}[ht]
	\centering
	\subfloat[][Step 1]{ 
		\fbox{\includegraphics[width=5.5cm]{../pictures/paper/ec_01_3.pdf}}
	    \label{fig:exec1}
	}
	\subfloat[][Step 2]{
		\fbox{\includegraphics[width=5.5cm]{../pictures/paper/ec_01_5.pdf}}
	    \label{fig:exec2}
	}
	\caption[]{Critical steps of a sequence for sentence
          (\ref{bsp:target-related}) where the early-closure reading
          (\ref{bsp:target-related-EC}) can be judged first.}
	\label{fig:exec}
\end{figure}
%
If participants adopted an early-closure reading as in
(\ref{bsp:target-related-EC}), a \emph{false}-judgment would be
expected at the position illustrated in \ref{fig:exec1}, as none of
the circles connected to the letter contain any suns. Under the
late-closure reading in (\ref{bsp:target-related-LC}), in contrast, a
\emph{true}-response would be expected no sooner than in the step
shown in Figure~\ref{fig:exec2}. The order of possible judgements is
reversed in the second kind of sequence including pictures like
\ref{fig:exlc1} followed by pictures like \ref{fig:exlc2}. Here, the
late-closure reading in (\ref{bsp:target-related-LC}) can be judged
first requiring a \emph{true}-response on \ref{fig:exlc1}. The
early-closure reading in (\ref{bsp:target-related-EC}) can only be
judged later in the sequence. When we reach \ref{fig:exlc2} a
\emph{false}-response is expected under an
early-closure/high-attachment reading.

@@

\begin{enumerate}[(i)]
\item a higher F0 maximum or a higher pitch range (difference 
between minimal and maximal F0 values), and
\item a longer duration of the accented element as opposed to its 
non-accented counterpart.
\end{enumerate}

\noindent We therefore recorded sentences by a trained speaker
familiar with the concept of contrastive focus and instructed her to
produce contrast as described in 1 and 2. Both \as-sentences and
\es-sentences were recorded in two versions each. In the accented
version, a pitch accent was placed on {\it einige}. The second version
had neutral prosody.  If accentuation is the driving force for local
readings, we would expect a higher proportion of local readings in the
former than in the latter version of the sentences.

Preference-related controls also differed prosodically. Modifier
attachment ambiguities comparable to the examples in
(\ref{bsp:target-related}) and (\ref{bsp:target-related2}) have been
shown to be sensitive to differences in prosodic phrasing. Though
there has been discussion as to whether speakers reliably
\textit{produce} disambiguating cues in such constructions
\citep[e.g.][]{Allbritton96,Kraljik05, Snedeker03, Schafer00}, overt
prosodic boundaries are generally acknowledged to guide interpretive
processes of late-closure sentences in \textit{comprehension}
\citep[e.g.][]{Steinhauer99, Augurzky06}.  Specifically, it has been
shown that prosodically separating a modifier from the directly
preceding material supports an early-closure reading, whereas
separating the two potential attachment sites is supportive of a late
closure reading. Thus, a phrase boundary between {\it squares} and
{\it with} in (\ref{bsp:target-related}) corresponds to a
high-attachment reading, whereas a prosodic phrase boundary following
{\it circles} corresponds to a low-attachment reading.

To test whether prosodic information is taken into account in our
task, we presented sentences in both of these prosodic variants. In
addition, in order to test prosody-independent preferences, we also
included a neutral version without any pronounced boundary. Further,
we also presented unambiguous sentences corresponding to the
early-closure reading. As in the case of the \as- and \es-sentences,
the unambiguous counterparts served as a baseline to control for
response biases that are independent of sentence meaning.

@@

We constructed a set of 15 items for German \as- and \es-sentences,
respectively, analogous to the examples in \ref{ex:as} and
\ref{ex:es}. Sentences were introduced by a subject {\small
  \acro{DP}}, including the quantifiers {\it alle diese} ("all of
these") or {\it genau eine(r) der} ("exactly one of these"), as well
as the head noun which denoted different icons, such as letters or
bells. Subjects were followed by an auxiliary, and a {\small
  \acro{PP}} containing {\it einige}, a possessive pronoun, and a noun
denoting a geometrical object, e.g. {\it einigen seiner
  Dreiecke/Quadrate} ("some of its triangles/squares"). The possessive
pronoun was used in order to fix relative quantifier scope to surface
scope, thus ensuring that {\it einige} has embedded status. The last
word was the main verb {\it verbunden} ("connected").  Each sentence
was recorded in a stressed and an unstressed version of scalar {\it
  einige}. As described above, three control
sentences were created for each experimental item, corresponding to
the critical positions in the course of uncovering the accompanying
sequences.  The positional controls were recorded with
neutral prosody.  Items were evenly distributed across 5 lists using a
Latin square design.

@@

For all target sentences (\as and \es), the determiner \emph{einige} was
produced with a contrastive pitch accent as well as with a neutral
accent. Contrastive accents in German are realized by an increase in F0 (range)
as well as by a longer duration (see Section~\ref{sec:design} above). A set of 15 
experimental items was recorded for each condition (\as vs.~\es, \emph{accented} 
vs.~\emph{unaccented}), resulting in a total number of 60 target sentences.

@@

\noindent Note that differences between Regions 1, 2, 4 and 7 can be expected
due to lexical differences between the \as and \es-conditions. As
preference-related fillers did not differ with respect to their lexical
properties, we carried out word-by-word analyses for these
conditions. Durational values included the respective word plus any
following silent interval. We did not include the
disambiguated fillers in these analyses as they involved very
different sentence types (i.e., constructions involving prepositional
phrases vs.~relative clauses).

@@

For the durational analyses, constituents were automatically labeled
by the \emph{Aligner} tool \citep{Rapp98}, and the
obtained values were manually corrected afterwards. For the targets,
two-factorial \acros{anova} with the factors \acro{Quantifier} (\emph{all}
vs.~\emph{exactly one}) and \acro{Prosody} (accented vs.~unaccented) were
carried out. For the preference-related fillers, we carried out
one-factorial \acros{anova} with the factor \acro{Prosody} (early
boundary, late boundary, neutral prosody). F0 values were extracted by
means of special Praat scripts
(\url{http://www.fon.hum.uva.nl/praat/}). For the present analyses,
differences between minimal and maximal F0 values for each region or
word were calculated. Again, two-factorial \acros{anova} were carried
out for statistical comparison of the target sentences, and one-factorial
\acros{anova} were carried out for preference-related controls.

The details of these analyses are found in
Appendix~\ref{sec:audit-sent-mater}. \dn{Petras Kommentar: Vielleicht
  sollten wir hier jeweils 1 Figure für akzentuiert vs. nicht und
  boundary vs. nicht reinbringen. Fabian meint zurecht, dass man nicht
  einfach die Ergebnisse nur in den Appendix packen kann. Es wären
  dann halt 4 Figures, da wir ja F0 und duration ansehen, und wir
  würden nur die experimentell interessanten veranschaulichen und die
  Statistik dann im Appendix lassen - was meint Ihr? --- Micha sagt ``Ja!''}

In sum, our speaker reliably
produced (i) differences in accent realization for the target
sentences and (ii) the expected boundary realizations for the
preference-related controls. Whereas accented elements clearly
differed from their unaccented counterpart by showing an increase in
duration and F0 range, prosodic boundaries for our preference-related
fillers were realized by pre-final lengthening (i.e., an increase in F0
and duration at the position preceding the boundary).

@@

For the preference-related controls, we used two types of
sequences. First, a sequence in which the crirical position for the
(preferred) late-closure reading was presented preceding the critical
position for the early-closure reading ({\it LC sequence}), and
second, a sequence in which the readings were presented in reversed
order ({\it EC sequence}. In the LC sequences, seven pictures were
presented, with the LC reading given at the third picture, and the EC
reading at the sixth picture. EC sequences consisted of five pictures,
in which the critical positions were realized at the third picture (EC
reading) and at the fifth picture (LC reading).

@@

In the following, we will report the results for targets and
preference-related controls separately. After providing the
descriptive results, we present statistical analyses, which were
carried out in the following manner.  We computed log-linear models
(see SCHEEPERS 2003 ????) in two variants (by item and by participant). In
the first, the factor {\it Item} was included, and in the second
analysis, the factor {\it Participants} was included. Inclusion of
these two factors allowed us to test whether the distribution of
participants' choices was identical accros items and participants. We
report log-likelihood ratio Chi-squares ($LRCS_1$ and $LRCS_2$),
degrees of freedom ($df_1$, $df_2$) and significance levels ($p_1$ and
$p_2$). The relevant experimental factors are provided in the specific
subsections below.

\paragraph{Target conditions and positional controls}
Performance on positional controls was high (92\% correct responses on
average), indicating that subjects understood the task, giving
truth-value judgements at the right moment in a sequence without being
influenced in general by inessential features of the picture material
or showing general response biases.

@@

      \begin{axis}[ybar, 
                  enlarge x limits=0.5,
                  enlarge y limits=0.25,
                  legend style={at={(0.5,-0.15)}, anchor=north, legend columns=-1},
        ylabel={\% of answers (rounded)}, symbolic x
        coords={neutral,accented}, xtick=data, nodes
        near coords, nodes near coords align={vertical},x=65, bar
        width=3mm, ymin = 8, ymax = 73 ]

@@

\subsection{Summary \& Interpretation}
\label{sec:interpretation}
